---
title: "SP2021coinoc"
author: "James C. Kosmopoulos"
date: "`r format(Sys.Date())`"
output: github_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r Load packages, message=FALSE, warning=FALSE}
# global options
knitr::opts_chunk$set(echo = TRUE)

# set working directory
setwd("~/Documents/Research/Rprojects/SP2021coinoc")

# load packages
library("knitr") # produce the knitted doc
library("gridExtra") # for arranging plots
library("cowplot")
```

# Load data

```{r}
data.raw <- read.csv("./data_raw.csv", header=TRUE)
data.raw$Inoc <- as.factor(data.raw$Inoc)

data <- data.raw %>% filter(Plant != "L-B6") %>% droplevels(.)
data <- data %>% filter(Plant != "L-C2") %>% droplevels(.)
data <- data %>% filter(Plant!= "R-A1") %>% droplevels(.)

data.long <- gather(data, key = "Trait", value = "Measurement", Tot_nod_mass:Nod_mass_per_root_mass)# use gather to stack the trait columns
```


# Sumamrize data

```{r}
library("tidyverse") # includes ggplot2, dplyr, readr, stringr

data.sum <- data %>%
  group_by(Inoc) %>%
  summarize(count = n(),
            # Tot_nod_mass
            mean_Tot_nod_mass = mean(Tot_nod_mass, na.rm = TRUE), sd_Tot_nod_mass = sd(Tot_nod_mass, na.rm = TRUE),
            SE_Tot_nod_mass = sd_Tot_nod_mass/sqrt(count),
            # Avg_nod_mass
            mean_Avg_nod_mass = mean(Avg_nod_mass, na.rm = TRUE), sd_Avg_nod_mass = sd(Avg_nod_mass, na.rm = TRUE),
            SE_Avg_nod_mass = sd_Avg_nod_mass/sqrt(count),
            # Root_mass
            mean_Root_mass = mean(Root_mass, na.rm = TRUE), sd_Root_mass = sd(Root_mass, na.rm = TRUE),
            SE_Root_mass = sd_Root_mass/sqrt(count),
            # Shoot_mass
            mean_Shoot_mass = mean(Shoot_mass, na.rm = TRUE), sd_Shoot_mass = sd(Shoot_mass, na.rm = TRUE),
            SE_Shoot_mass = sd_Shoot_mass/sqrt(count),
            # Shoot_mass_per_root_mass
            mean_Shoot_mass_per_root_mass = mean(Shoot_mass_per_root_mass, na.rm = TRUE), sd_Shoot_mass_per_root_mass = sd(Shoot_mass_per_root_mass, na.rm = TRUE),
            SE_Shoot_mass_per_root_mass = sd_Shoot_mass_per_root_mass/sqrt(count),
            # Shoot_mass_per_nod_mass
            mean_Shoot_mass_per_nod_mass = mean(Shoot_mass_per_nod_mass, na.rm = TRUE), sd_Shoot_mass_per_nod_mass = sd(Shoot_mass_per_nod_mass, na.rm = TRUE),
            SE_Shoot_mass_per_nod_mass = sd_Shoot_mass_per_nod_mass/sqrt(count),
            # Nod_mass_per_root_mass
            mean_Nod_mass_per_root_mass = mean(Nod_mass_per_root_mass, na.rm = TRUE), sd_Nod_mass_per_root_mass = sd(Nod_mass_per_root_mass, na.rm = TRUE),
            SE_Nod_mass_per_root_mass = sd_Nod_mass_per_root_mass/sqrt(count)
            )

data.long.sum <- data.long %>%
  group_by(Inoc, Trait) %>%
  summarize(count = n(),
            mean_Trait = mean(Measurement, na.rm = TRUE), sd_Trait = sd(Measurement, na.rm = TRUE),
            SE_Trait = sd_Trait/sqrt(count))           
data.long.sum$Trait <- factor(data.long.sum$Trait, level = c("Tot_nod_mass", "Nod_num", "Avg_nod_mass", "Root_mass", "Shoot_mass", "Shoot_mass_per_root_mass", "Shoot_mass_per_nod_mass", "Nod_mass_per_root_mass"))
```



# Shapiro tests

Test if the trait values are normally distributed. If p<0.05, then the data range is not normally distrubuted.

## Tot_nod_mass
```{r}
shapiro.test(data$Tot_nod_mass) #p<0.05, not normal
hist(data$Tot_nod_mass) # the data are heavily left-skewed
hist(sqrt(data$Tot_nod_mass)) # transform
shapiro.test(sqrt(data$Tot_nod_mass)) #p>0.05, transformed data are normal

```

## Nod_num
```{r}
shapiro.test(data$Nod_num) #p<0.05, not normal
hist(data$Nod_num) # the data are heavily left-skewed
hist(sqrt(data$Nod_num)) # transform
shapiro.test(sqrt(data$Nod_num)) #p>0.05, transformed data are normal

```

## Avg_nod_mass
```{r}
shapiro.test(data$Avg_nod_mass) #p>0.05, normal
hist(data$Avg_nod_mass) # looks normal

```

## Root_mass
```{r}
shapiro.test(data$Root_mass) #p>0.05, normal
hist(data$Root_mass) # looks normal

```

## Shoot_mass
```{r}
shapiro.test(data$Shoot_mass) #p>0.05, normal
hist(data$Shoot_mass) # looks normal

```

## Shoot_mass_per_root_mass
```{r}
shapiro.test(data$Shoot_mass_per_root_mass) #p<0.05, not normal
hist(data$Shoot_mass_per_root_mass) # slightly left skewed

```

## Shoot_mass_per_nod_mass
```{r}
shapiro.test(data$Shoot_mass_per_nod_mass) #p<0.05, not normal
hist(data$Shoot_mass_per_nod_mass) # heavily left skewed

```

## Nod_mass_per_root_mass
```{r}
shapiro.test(data$Nod_mass_per_root_mass) #p<0.05, not normal
hist(data$Nod_mass_per_root_mass) # heavily left skewed

```


# Linear mixed effects model
Build a function with a linear mixed effects model. When estimated marginal means are calculated for the model, it can be used to test for significant differences in the means of each trait when making distinct pairwise comparisons. NOTE: This assumes normality in the data being tested, and I already know that not all of the data (tot_nod_mass and the ratios) are normally distributed, so this might not be the right test for all of the traits. Furthermore, this is also probably not the right statistical test to use for my ratios (like nod_mass_per_root_mass), and I need some other test for that.

```{r}
library("emmeans") # for posthoc data analysis
library("lme4") # linear mixed models
library("car") # for Anova function

function.lmer <- function(trait) {
  lmer.fill.bench <- "+ (1|Bench)"
  lmer.fill <- paste(trait, "~", "Inoc", lmer.fill.bench, sep = " ", collapse = NULL)
  lmer <- lmer(lmer.fill, data = data)
  print(summary(lmer))
  
  anova.lmer <-  Anova(lmer, type=2)
  print(anova.lmer)
  
  lmer.emms <- emmeans(lmer, pairwise ~ Inoc)
  lmer.emms.sum <- summary(lmer.emms, infer = c(TRUE,TRUE), adjust = "fdr")
  print(lmer.emms.sum)
  emm.out.path <- paste("./emmeans", trait, sep="/", collapse=NULL)
  emm.out.path.csv <- paste(emm.out.path, "csv", sep=".", collapse = NULL)
  write.csv(lmer.emms.sum$contrasts, file=emm.out.path.csv, row.names = FALSE)
  
  plot(lmer)
}

function.lmer.alltraits <- function() {
  traitlist = c("Tot_nod_mass", "Nod_num", "Avg_nod_mass", "Root_mass", "Shoot_mass", "Shoot_mass_per_root_mass", "Shoot_mass_per_nod_mass", "Nod_mass_per_root_mass")
  for(i in traitlist){
    function.lmer(i)
  }
}
```


